---
title: "Lesson 8"
author: "Team Rython, Dainius Masiliunas and Tim Weerman"
date: "January 14, 2016"
output: html_document
---

Apache License 2.0 

***

####Needed packages
```{r message = FALSE}
library(raster)
```

***

####Source
We start with two source lines and right after them the code from those two scripts.
```{r eval = TRUE}
source("src/RMSE.r")

```
Calculate the Root Mean Squared Error.
```{r eval = FALSE}

RMSE = function(truth, prediction)
{
    return(sqrt(mean((truth-prediction)^2, na.rm=TRUE)))
}
```

Calculate partial RMSE for different zones by generating squared difference rasters.
```{r eval = FALSE}
GetDifferenceRaster = function(former, latter, filename=paste("data/", names(latter), ".grd", sep=""), force=FALSE)
{
    # Figure out a reasonable output name
    if (filename == "data/layer.grd")
        warning("Please pass filename or assign meaningful names to input rasters! Else the loaded data may not match expectations.")
    
    # Create one or load one
    if (!file.exists(filename) || force)
    {
        return(overlay(former, latter, fun=function(truth, prediction){(truth-prediction)^2},
            filename=filename))
    }
    else
        return(raster(filename))
}
```
Calculate RMSE per zone.
```{r}
StratifiedRMSE = function(truth, prediction, zones, zonenames = "", ...)
{
    # Unfortunately, zonal() just passes a vector of numbers, not a matrix.
    # So we have to calclate RMSE manually from a difference raster.
    differenceRaster = GetDifferenceRaster(truth, prediction, ...)
    
    # Get a mean from the zones in the difference raster
    zonestats = zonal(differenceRaster, zones, fun="mean")
    
    # Add nice labels, if we have them
    if (length(zonenames) > 1)
        rownames(zonestats) = zonenames
        
    return(zonestats)
}
```

***

The second source file starts here.
```{r eval = TRUE}
source("src/LinearModel.r")
```
Creates a linear model and prints its summary, and if desired whether some covariates ought to be removed from the model, and a plot of residuals.
```{r eval = FALSE}
lmValidation = function(..., printstep=FALSE, printplot=FALSE)
{
    LM = lm(...)
    print(summary(LM))
    if (printstep)
        step(LM)
    if (printplot)
        plot(LM)
    return(LM)
}

# Creates a prediction based on the model, clamps values to a given range, and
# if requested plots a histogram and/or a comparison plot between the layers
predictValidation = function(object, ..., truthlayer = "", range = c(-Inf, +Inf), plothist = FALSE, plotcomparison = FALSE)
{
    Prediction = predict(object, ...)
    Prediction[Prediction < range[1]] = NA
    Prediction[Prediction > range[2]] = NA
    if (plothist)
        hist(Prediction, breaks = 200, main="Histogram of predicted values")
    if (plotcomparison)
    {
        TruthExists = try(class(object[[truthlayer]]))
        if (class(TruthExists) == "try-error")
        {
            warning("Requested a comparison plot, but could not find a layer to compare to!")
            return(Prediction)
        }
        op = par(mfrow=c(1,2))
        plot(Prediction, colNA="black", main="Predicted values")
        plot(object[[truthlayer]], colNA="black", main="Ground truth")
        par(op)
    }
    return(Prediction)
}
```

***

####Download/load information
```{r eval= TRUE}
GetFromWURgit = function(filename)
{
    download.file(paste("https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/raw/gh-pages/data/", filename, sep=""),
        paste("data/", filename, sep=""), "wget")
    return(paste("data/", filename, sep=""))
}
load(GetFromWURgit("GewataB1.rda"))
load(GetFromWURgit("GewataB2.rda"))
load(GetFromWURgit("GewataB3.rda"))
load(GetFromWURgit("GewataB4.rda"))
load(GetFromWURgit("GewataB5.rda"))
load(GetFromWURgit("GewataB7.rda"))
load(GetFromWURgit("vcfGewata.rda"))
load(GetFromWURgit("trainingPoly.rda"))  

DataBrick = brick(GewataB1, GewataB2, GewataB3, GewataB4, GewataB5, GewataB7, vcfGewata)
names(DataBrick) = c("Blue", "Green", "Red", "NIR", "SWIR", "Emission", "VCF")
```

Sanitise data
```{r eval= TRUE }
DataBrick[["VCF"]][DataBrick[["VCF"]] > 100] = NA
DataValues = as.data.frame(getValues(DataBrick))
```
Create a training raster for RMSE
```{r message = FALSE}
trainingRaster = rasterize(trainingPoly, DataBrick[["VCF"]], field='Class', filename="data/trainingRaster.grd", overwrite=TRUE)
```

***

Here we produce plots that demonstrate the relationship between the Landsat bands and the VCF tree cover. 
```{r eval = TRUE}
pairs(DataBrick)
```

From this we can conclude that they are all negatively correlated with VCF, except for NIR. The bands are also highly correlated with each other.

***


Here we create a lm() model and show a summary with our own lmValidation()
```{r eval= TRUE}
LM = lmValidation(VCF ~ Blue + Green + Red + NIR + SWIR + Emission, data=DataValues, printstep = TRUE)
```
Emission can be dropped
```{r eval= TRUE}
LM = lmValidation(VCF ~ Blue + Green + Red + NIR + SWIR, data=DataValues, printstep = TRUE)
```
Nothing further can be dropped, the most significant bands are NIR and green.  
NB: a linear model isn't very appropriate, because normality and independence assumptions are violated!  


```{r eval = TRUE}
BrickSubset = dropLayer(DataBrick, "Emission")
```

***

Here we  predict the values and plot a histogram using the predictValidation().
```{r eval = TRUE}
Prediction = predictValidation(BrickSubset, model=LM, na.rm=TRUE, plothist=TRUE)
```

In the histogram we see that some values are out of range, we apply range constraints to the prediction analysis.  
The results are plotted below.
```{r eval = TRUE}
Prediction = predictValidation(BrickSubset, model=LM, na.rm=TRUE, overwrite=TRUE, filename="data/PredictionRaster.grd",
    truthlayer="VCF", range=c(0, +Inf), plotcomparison=TRUE)
names(Prediction) = "Predicted.LM"
```

Compute the RMSE:
```{r eval = TRUE}
RMSE(getValues(DataBrick[["VCF"]]), getValues(Prediction))
```

RMSE per zone:
```{r eval = TRUE}
zonestats = StratifiedRMSE(DataBrick[["VCF"]], Prediction, trainingRaster, zonenames=levels(trainingPoly@data$Class))
zonestats
```

As a bonus, plot the difference raster:
```{r eval = TRUE}
plot(raster("data/Predicted.LM.grd"))
```

***

Here we repeat the analysis and only use independent variables:
```{r eval = TRUE}
ReducedBrick = dropLayer(DataBrick, "Emission")
ReducedBrick = dropLayer(ReducedBrick, "Green")
ReducedBrick = dropLayer(ReducedBrick, "Red")
ReducedBrick = dropLayer(ReducedBrick, "SWIR")

pairs(ReducedBrick)
ReducedLM = lmValidation(VCF ~ Blue + NIR, data=DataValues, printstep=TRUE)
RedPrediction = predictValidation(ReducedBrick, model=ReducedLM, na.rm=TRUE,
    filename="data/ReducedPredictionRaster.grd", overwrite=TRUE, range=c(0, +Inf),
    truthlayer="VCF", plothist=TRUE, plotcomparison=TRUE)
names(RedPrediction) = "Predicted.LM.Blue.NIR"
RMSE(getValues(ReducedBrick[["VCF"]]), getValues(RedPrediction))
zonestatsR = StratifiedRMSE(DataBrick[["VCF"]], RedPrediction, trainingRaster,
    zonenames=levels(trainingPoly@data$Class))
zonestatsR
plot(raster("data/Predicted.LM.Blue.NIR.grd"))
```

***
